Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2018-12-09 manuscript.tex
first draft

\shorttitle{Contextual recruitment of selective attention}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} The Graduate Center of the City University of New York\\\textsuperscript{2} Brooklyn College of the City University of New York}
\keywords{Attention, Memory, Cognitive Control, Conflict Adaptation, Context-specific, Task Relevance\newline\indent Word count: 5946}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

@@

\section{Introduction}\label{introduction}

Selective attention processes are commonly investigated using
interference paradigms such as the Stroop (1935) and flanker (B. A.
Eriksen \& Eriksen, 1974) tasks, where participants identify a target
while ignoring a response-congruent or -incongruent distractor.
Performance is typically better on congruent versus incongruent trials
and the difference-the congruency effect-taken as an index of
attentional priorities. Large congruency effects are thought to reflect
ineffective filtering of the distracting stimuli whereas small
congruency effects are thought to reflect effective filtering. By
probing factors that systematically alter congruency effects, we can
then make inferences about processes that control attentional filtering.
For example, manipulating the frequency of conflict via the proportion
of congruent versus incongruent trials has shown to influence the size
of the congruency effect. Typically, a high proportion congruent
experiment produces large congruency effects, whereas a low proportion
congruent experiment produces small congruency effects (Logan \&
Zbrodoff, 1979; Lowe \& Mitterer, 1982; West \& Baylis, 1998). This
result is usually explained as strategic control, where participants
increase attentional control under high-conflict demands and relax
attentional control under low-conflict demands (Logan, 1980; Logan \&
Zbrodoff, 1979; Logan, Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer,
1982). Recent work however, has demonstrated that attentional control is
not only adjusted by top-down regulation, but can also be triggered
automatically by environmental cues (Brosowsky \& Crump, 2018; J. Bugg
\& Crump, 2012; Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \&
Egner, 2012; Mayr \& Bryck, 2007).

For example, Crump, Gong, and Milliken (see also, Corballis \& Gratton,
2003; Crump, Gong, \& Milliken, 2006) presented Stroop stimuli in one of
two randomly chosen locations and manipulated the frequency of conflict
associated with each location. One location was associated with a high
frequency of conflict (25\% congruent trials) and the other with a low
frequency of conflict (75\% congruent trials). Overall, the proportion
of congruent trials was 50\% and randomized such that the upcoming
location could not be predicted. Even so, congruency effects were shown
to be smaller for trials where the stimulus appeared in the high
conflict location as compared to the low conflict location. This effect,
known now as the context-specific proportion congruent effect (CSPC),
has now been replicated in a number of different selective attention
paradigms (e.g., Alards-Tomalin, Brosowsky, \& Mondor, 2017; Blais,
Harris, Sinanian, \& Bunge, 2015; J. Bugg, 2014; Crump, 2016; Crump,
Milliken, Leboe-McGowan, Leboe-McGowan, \& Gao, 2018; Fischer,
Gottschalk, \& Dreisbach, 2014; Hübner \& Mishra, 2016).

Critical evidence however, that CSPC effects reflect context-specific
control rather than other non-control learning processes (e.g., Schmidt
\& Besner, 2008), comes from work showing that CSPC effects can transfer
to frequency unbiased items (Brosowsky \& Crump, 2016; Crump \&
Milliken, 2009; though, see Hutcheon \& Spieler, 2017; Weidler \& Bugg,
2016; Weidler, Dey, \& Bugg, 2018). Crump and Milliken (2009), for
example, divided Stroop items into two mutually exclusive sets (e.g.,
RED/GREEN and BLUE/YELLOW). One set was defined as the frequency biased
set, and presented with 75\% congruency in one location, and 25\%
congruency in the other. The second set however, was presented with 50\%
congruency in both locations. Nevertheless, they found smaller
congruency effects for unbiased items presented in the high conflict
location as compared to the low conflict location.

@@

Task-relevance might be one factor that is important for establishing
and using context-attention associations. This proposal might seem
counter-intuitive as it is typically assumed that the context, within
selective attention tasks at least, is completely irrelevant to the
task. In fact, referring to such stimulus dimensions as the
\enquote{context} carries with it an implication that the feature is
background information, irrelevant to the task-at-hand, and should be
ignored by the participant. A closer look at CSPC designs however,
suggests that the context cues may not be as irrelevant as implied. For
example, the most commonly used context cue is location (Brosowsky \&
Crump, 2016; Corballis \& Gratton, 2003; Crump, 2016; Crump et al.,
2006; Crump, Vaquero, \& Milliken, 2008; Weidler \& Bugg, 2016; Weidler
et al., 2018). Although in some sense, the location is irrelevant for
identifying the color of a word, it seems disingenuous to consider it
completely task-irrelevant. In order to identify the target, the
participant must first identify and orient to its location on the
screen. In this is sense, it is relevant to their ability to complete
the task and furthermore, location information may receive priority
during encoding (Logan, 1998; Mayr, 1996). Similar arguments could be
made about other context features that are inherent to processing the
target stimulus, like font-type (J. Bugg, Jacoby, \& Toth, 2008) and
color (Vietze \& Wendt, 2009).

To date, we are aware of only one study that has looked at
task-relevance within the CSPC literature. Crump, Vaquero, \& Milliken
(2008) used a Stroop prime-probe task where the distractor word is
presented first followed by a color patch probe. Critically, they used
the shape of the color patch probe, either a square or a circle, as the
contextual dimension; one associated with a high frequency of conflict,
the other with a low frequency. In one experiment participants were made
aware of the shape-proportion contingencies but given no additional
instructions. In the second, participants were given a secondary task to
count the number of trials that contained a square. The first experiment
failed to find evidence for context-specificity, whereas the second did.
Although not testing task-relevance directly, (2013)) found similar
effects. In this study they used images of male and female faces as
context cues in a flanker task. When given instructions to think of the
faces as members of the gender categories they found CSPC effects, but
when given instructions to think of the faces as individuals they did
not. These results suggest that making a context dimension relevant to
the on-going task is important for producing CSPC effects.

\section{The Current Study}\label{the-current-study}

Some prior suggests that the task-relevance of contextual information is
important for establishing new associations between context cues and
attentional priorities. Simply put, CSPC effects emerge when context
cues are somehow made relevant to the task, but do not when made
irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, it
still remains unclear how task-relevance may or may not inform the use
of previously learned associations. In the current study, we examined
whether manipulating task-relevance would cause participants to ignore a
previously learned association and adopt a new context-specific control
strategy.

A flanker task was used to measure attentional control in an adapted
CSPC design (Crump \& Milliken, 2009; Crump, Brosowsky, \& Milliken,
2017). Similar to previous studies, flanker stimuli were presented in
contexts associated with differing levels of conflict (0\%, 50\%, or
100\% congruent). However, unlike previous CSPC designs where contexts
were defined by a single discriminating feature (e.g., upper vs.~lower
screen locations), contexts were defined by two feature dimensions (see
Figure 1A). In one condition, for example, contexts were defined by
object identity (hat or chair) and color (blue or green). Only three out
of the four possible feature combinations were presented, each
associated with a different proportion of congruent trials (0\%, 50\%,
or 100\% congruent). Critically, the frequency unbiased context (50\%
congruent) always shared a feature with each of the frequency-biased
contexts (0\% and 100\% congruent). The frequency biased contexts (0\%
and 100\% congruent) however, did not share any features (see Figure 1C
for an example).

@@

The critical measure of interest is the congruency effect produced in
the frequency unbiased context. If, on the one hand, task-relevance has
no impact on the use of context-attention associations we would expect
no differences between the congruency effects when the high-conflict
context dimension is made task-relevant or the low-conflict context
dimension is made task-relevant. This could occur because all three
contexts are treated as individual contexts throughout the whole
experiment (e.g., Cañadas et al., 2013). Or it could occur because
associations formed in the first block interfere with learning new
associations in the second (e.g., Brosowsky \& Crump, 2016). On other
hand, task-relevance might not only dictate how associations are formed,
but also whether learned associations are used. In this case, we would
expect that there will be smaller congruency effects when the
high-conflict context dimension is made task-relevant as compared to
when the low-conflict context dimension is made task-relevant. This
would demonstrate that participants were able to ignore a previously
learned association and adopt a new context-specific control strategy.

A secondary goal was to conceptually replicate previous findings and
test the generalizability of any task-relevance effects. Therefore, we
included three conditions that differed only in the kinds of stimuli
used to create context cues. In one condition we used face images where
context dimensions were defined by social categories, similar to Cañadas
et al., (2013). However, they found that social categorization occurred
spontaneously and it was unclear what impact that would have on the
current results. Therefore, we included another condition that used
images of simple objects where context dimensions were defined by object
features. Finally, Cañadas et al.(2013) also found that context effects
generalized to novel face images. In our third condition we used
non-repeating face exemplars to examine whether any task-relevance
effects would also generalize in our design or if learning would be
image-specific.

\section{Methods}\label{methods}

@@

\subsection{Participants}\label{participants}

All participants were recruited from Amazon Mechanical Turk (AMT) and
compensated \$2.00 for participating. The amount compensated was
calculated by estimating the maximum amount of time required to complete
each experiment and multiplying by \$6.00 per hour. For each experiment
the number of HITs (Human intelligence tasks, an Amazon term for a
work-unit) refers to the number of participants who initiated the study.
Participants were included in the study if they completed all trials. We
posted 150 HITs and 145 participants completed all trials.

\subsection{Apparatus \& Stimuli}\label{apparatus-stimuli}

@@

\subsection{Procedure}\label{procedure}

@@

\section{Results}\label{results}

\begin{figure}
\centering
\includegraphics{v1-task-relevance-cspc_files/figure-latex/figure2-1.pdf}
\caption{\label{fig:figure2}Results from experiment 1 showing congruency effects
(incongruent - congruent) for frequency unbiased items in reaction times
(left) and error rates (right) as a function of the task-relevant
context (100\% proportion congruent versus 0\% proportion congruent).}
\end{figure}

@@

Similarly, the corresponding error analysis also resulted in a
significant two-way interaction between the task-relevant context and
unbiased-item congruency, \(F(1, 128) = 8.30\),
\(\mathit{MSE} = 1,608.54\), \(p = .005\), \(\hat{\eta}^2_p = .061\),
and a non-significant three-way interaction between task-relevant
context, unbiased-item congruency, and context-type,
\(F(2, 128) = 0.00\), \(\mathit{MSE} = 1,608.54\), \(p = .996\),
\(\hat{\eta}^2_p = .000\). Therefore, the congruency effects in error
rates were also smaller when the context dimension associated with high
conflict was made task-relevant, corroborating the results of the
reaction time analysis.

\section{General Discussion}\label{general-discussion}

The aim of the current study was to investigate whether task-relevance
plays a role in the contextual recruitment of selective attention.
Specifically, we tested whether manipulating the relative task-relevance
of context cues could cause participants to ignore a previously learned
context-association and apply a new association. We created a frequency
unbiased context cue that shared features with two frequency biased
contexts and included a feature-counting task to manipulate the
task-relevance of context dimensions across two blocks of trials.
Critically, halfway through the experiment participants received new
instructions changing the task-relevant feature from one frequency
biased cue to the other.

@@

In light of prior work, we take this as evidence that the contextual
recruitment of selective attention, although likely implicit, is not
obligatory (e.g., Brosowsky \& Crump, 2016), requiring that
environmental information be incorporated into the task representation.
Similarly, there appears to be flexibility in which environmental
features are selected and used to guide attention, which can be rapidly
updated depending on the task-relevance of those features. Such a result
lends some insight into how context-specific control might operate in
more complex, real-world environments, where there is an over-abundance
of environmental features that afford many different learned
associations. From a theoretical perspective, this result is consistent
memory-based accounts of CSPC phenomena (Brosowsky \& Crump, 2018; J.
Bugg \& Hutchison, 2013; Crump \& Milliken, 2009; Crump et al., 2018,
2017). Under this view, a memory process encodes attentional priorities
in the representation of individual experiences and, as a result, become
associated with the environment where they were used. The subsequent
reoccurrence of a prior context triggers the retrieval and reinstatement
of those attentional priorities. Our results show however, that all the
features of the environment may not be treated equally and that only
task-relevant features are used to probe memory and guide attention.

@@

\section{References}\label{references}

@@

\hypertarget{refs}{}
\hypertarget{ref-abrahamse_grounding_2016}{}
Abrahamse, E., Braem, S., Notebaert, W., \& Verguts, T. (2016).
Grounding cognitive control in associative learning. \emph{Psychological
Bulletin}, \emph{142}, 693--728.
doi:\href{https://doi.org/10.1037/bul0000047}{10.1037/bul0000047}

\hypertarget{ref-alards-tomalin_auditory_2017}{}
Alards-Tomalin, D., Brosowsky, N. P., \& Mondor, T. A. (2017). Auditory
statistical learning: Predictive frequency information affects the
deployment of contextually mediated attentional resources on perceptual
tasks. \emph{Journal of Cognitive Psychology}, \emph{0}(0), 1--11.
doi:\href{https://doi.org/10.1080/20445911.2017.1353518}{10.1080/20445911.2017.1353518}

\hypertarget{ref-blais_trial-by-trial_2015}{}
Blais, C., Harris, M. B., Sinanian, M. H., \& Bunge, S. A. (2015).
Trial-by-trial adjustments in control triggered by incidentally encoded
semantic cues. \emph{The Quarterly Journal of Experimental Psychology},
\emph{68}, 1920--1930.
doi:\href{https://doi.org/10.1080/17470218.2014.1000346}{10.1080/17470218.2014.1000346}

\hypertarget{ref-brady_visual_2013}{}
Brady, T. F., Konkle, T., Gill, J., Oliva, A., \& Alvarez, G. A. (2013).
Visual long-term memory has the same limit on fidelity as visual working
memory. \emph{Psychological Science}, \emph{24}, 981--990.
doi:\href{https://doi.org/10.1177/0956797612465439}{10.1177/0956797612465439}

\hypertarget{ref-brewer_dual_1988}{}
Brewer, M. B. (1988). A dual process model of impression formation. In
R. S. Wyer Jr \& T. K. Srull (Eds.), \emph{A dual-process model of
impression formation: Advances in social cognition} (Vol. 1, pp. 1--36).
Hillsdale, NJ: Erlbaum.

\hypertarget{ref-brosowsky_context-specific_2016}{}
Brosowsky, N. P., \& Crump, M. J. (2016). Context-specific attentional
sampling: Intentional control as a pre-requisite for contextual control.
\emph{Consciousness and Cognition}, \emph{44}, 146--160.
doi:\href{https://doi.org/10.1016/j.concog.2016.07.001}{10.1016/j.concog.2016.07.001}

\hypertarget{ref-brosowsky_memory-guided_2018}{}
Brosowsky, N. P., \& Crump, M. J. C. (2018). Memory-guided selective
attention: Single experiences with conflict have long-lasting effects on
cognitive control. \emph{Journal of Experimental Psychology: General}.
doi:\href{https://doi.org/10.1037/xge0000431}{10.1037/xge0000431}

\hypertarget{ref-bugg_conflict-triggered_2014}{}
Bugg, J. (2014). Conflict-triggered top-down control: Default mode, last
resort, or no such thing? \emph{Journal of Experimental Psychology:
Learning, Memory, and Cognition}, \emph{40}(2), 567.

\hypertarget{ref-bugg_support_2012}{}
Bugg, J., \& Crump, M. J. (2012). In support of a distinction between
voluntary and stimulus-driven control: A review of the literature on
proportion congruent effects. \emph{Frontiers in Psychology}, \emph{3},
367.
doi:\href{https://doi.org/10.3389/fpsyg.2012.00367}{10.3389/fpsyg.2012.00367}

\hypertarget{ref-bugg_converging_2013}{}
Bugg, J., \& Hutchison, K. A. (2013). Converging evidence for control of
color--word Stroop interference at the item level. \emph{Journal of
Experimental Psychology: Human Perception and Performance},
\emph{39}(2), 433.

\hypertarget{ref-bugg_multiple_2008}{}
Bugg, J., Jacoby, L. L., \& Toth, J. P. (2008). Multiple levels of
control in the Stroop task. \emph{Memory \& Cognition}, \emph{36},
1484--1494.
doi:\href{https://doi.org/10.3758/MC.36.8.1484}{10.3758/MC.36.8.1484}

\hypertarget{ref-canadas_social_2013}{}
Cañadas, E., Rodríguez-Bailón, R., Milliken, B., \& Lupiáñez, J. (2013).
Social categories as a context for the allocation of attentional
control. \emph{Journal of Experimental Psychology: General}, \emph{142},
934--943. doi:\href{https://doi.org/10.1037/a0029794}{10.1037/a0029794}

\hypertarget{ref-corballis_independent_2003}{}
Corballis, P. M., \& Gratton, G. (2003). Independent control of
processing strategies for different locations in the visual field.
\emph{Biological Psychology}, \emph{64}, 191--209.
doi:\href{https://doi.org/10.1016/S0301-0511(03)00109-1}{10.1016/S0301-0511(03)00109-1}

\hypertarget{ref-crisp_multiple_2007}{}
Crisp, R. J., \& Hewstone, M. (2007). Multiple social categorization.
\emph{Advances in Experimental Social Psychology}, \emph{39}, 163--254.

>>>>

\shorttitle{Contextual recruitment of selective attention}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} The Graduate Center of the City University of New York\\\textsuperscript{2} Brooklyn College of the City University of New York}
\keywords{Attention, Memory, Cognitive Control, Conflict Adaptation, Context-specific, Task Relevance\newline\indent Word count: 4402}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}\singlespacing
\newenvironment{zeroindent} {\par\setlength{\parindent}{0pt}} {\par}
\setlength{\parskip}{0.25cm plus4mm minus3mm}

@@

Selective attention is commonly investigated using interference paradigms such as the Stroop (1935) and flanker (Eriksen \& Eriksen, 1974) tasks, where participants identify a target while ignoring a response-congruent or -incongruent distractor. Performance is typically better on congruent versus incongruent trials and the difference-the congruency effect-taken as an index of attentional priorities. Large congruency effects are thought to reflect ineffective filtering of the distracting stimuli whereas small congruency effects are thought to reflect effective filtering. By probing factors that systematically alter congruency effects, we can then make inferences about processes that control attentional filtering. For example, manipulating the frequency of conflict via the proportion of congruent versus incongruent trials has shown to influence the size of the congruency effect. Typically, a high proportion congruent experiment produces large congruency effects, whereas a low proportion congruent experiment produces small congruency effects (Logan \& Zbrodoff, 1979; Lowe \& Mitterer, 1982; West \& Baylis, 1998). This result is usually explained as strategic control, where participants increase attentional control under high-conflict demands and relax attentional control under low-conflict demands (Logan, 1980; Logan \& Zbrodoff, 1979; Logan, Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer, 1982). Recent work however, has demonstrated that attentional control is not only adjusted by top-down regulation, but can also be triggered automatically by environmental cues (Brosowsky \& Crump, 2018; Bugg \& Crump, 2012; Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \& Egner, 2012; Mayr \& Bryck, 2007).

For example, Crump, Gong, and Milliken (Crump, Gong, \& Milliken, 2006; see also, Corballis \& Gratton, 2003) presented Stroop stimuli in one of two randomly chosen locations and manipulated the frequency of conflict associated with each location. One location was associated with a high frequency of conflict (25\% congruent trials) and the other with a low frequency of conflict (75\% congruent trials). Overall, the proportion of congruent trials was 50\% and randomized such that the upcoming location could not be predicted. Even so, congruency effects were shown to be smaller for trials where the stimulus appeared in the high conflict location as compared to the low conflict location. This effect, known now as the context-specific proportion congruent effect (CSPC), has now been replicated in a number of different selective attention paradigms (e.g., Alards-Tomalin, Brosowsky, \& Mondor, 2017; Blais, Harris, Sinanian, \& Bunge, 2015; Bugg, 2014; Crump, 2016; Crump, Milliken, Leboe-McGowan, Leboe-McGowan, \& Gao, 2018; Fischer, Gottschalk, \& Dreisbach, 2014; Hübner \& Mishra, 2016).

Critical evidence however, that CSPC effects reflect context-specific control rather than other non-control learning processes (e.g., Schmidt \& Besner, 2008), comes from work showing that CSPC effects can transfer to frequency unbiased items (Brosowsky \& Crump, 2016; Crump \& Milliken, 2009; Weidler \& Bugg, 2016; Weidler, Dey, \& Bugg, 2018; though, see Hutcheon \& Spieler, 2017). Crump and Milliken (2009), for example, divided Stroop items into two mutually exclusive sets (e.g., RED/GREEN and BLUE/YELLOW). One set was defined as the frequency biased set, and presented with 75\% congruency in one location, and 25\% congruency in the other. The second set however, was presented with 50\% congruency in both locations. Nevertheless, they found smaller congruency effects for unbiased items presented in the high conflict location as compared to the low conflict location.

@@

Task-relevance might be one factor that is important for establishing and using context-attention associations. This proposal might seem counter-intuitive as it is typically assumed that the context, within selective attention tasks at least, is completely irrelevant to the task. In fact, referring to such stimulus dimensions as the \enquote{context} carries with it an implication that the feature is background information, irrelevant to the task-at-hand, and should be ignored by the participant. A closer look at CSPC designs however, suggests that the context cues may not be as irrelevant as implied. For example, the most commonly used context cue is location (Brosowsky \& Crump, 2016; Corballis \& Gratton, 2003; Crump, 2016; Crump et al., 2006; Crump, Vaquero, \& Milliken, 2008; Weidler \& Bugg, 2016; Weidler et al., 2018). Although in some sense, the location is irrelevant for identifying the color of a word, it seems disingenuous to consider it completely task-irrelevant. In order to identify the target, the participant must first identify and orient to its location on the screen. In this is sense, it is relevant to their ability to complete the task and furthermore, location information may receive priority during encoding (Logan, 1998; Mayr, 1996). Similar arguments could be made about other context features that are inherent to processing the target stimulus, like font-type (Bugg, Jacoby, \& Toth, 2008) and color (Vietze \& Wendt, 2009).

To date, we are aware of only one study that has looked at task-relevance within the CSPC literature. Crump, Vaquero, \& Milliken (2008) used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, Cañadas, Rodríguez-Bailón, Milliken, and Lupiáñez (2013){]} found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for producing CSPC effects.

\hypertarget{the-current-study}{%
\section{The Current Study}\label{the-current-study}}

Prior work suggests that the task-relevance of contextual information is important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects emerge when context cues are somehow made relevant to the task, but do not when made irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, it still remains unclear how task-relevance may or may not inform the use of previously learned associations. In the current study, we examined whether manipulating task-relevance would cause participants to ignore a previously learned association and adopt a new context-specific control strategy.

A flanker task was used to measure attentional control in an adapted CSPC design (Crump, Brosowsky, \& Milliken, 2017; Crump \& Milliken, 2009). Similar to previous studies, flanker stimuli were presented in contexts associated with differing levels of conflict (0\%, 50\%, or 100\% congruent). However, unlike previous CSPC designs where contexts were defined by a single discriminating feature (e.g., upper vs.~lower screen locations), contexts were defined by two feature dimensions (see Figure 1A). In one condition, for example, contexts were defined by object identity (hat or chair) and color (blue or green). Only three out of the four possible feature combinations were presented, each associated with a different proportion of congruent trials (0\%, 50\%, or 100\% congruent). Critically, the frequency unbiased context (50\% congruent) always shared a feature with each of the frequency-biased contexts (0\% and 100\% congruent). The frequency biased contexts (0\% and 100\% congruent) however, did not share any features (see Figure 1C for an example).

@@

The critical measure of interest is the congruency effect produced in the frequency unbiased context. If, on the one hand, task-relevance has no impact on the use of context-attention associations we would expect no differences between the congruency effects when the high-conflict context dimension is made task-relevant or the low-conflict context dimension is made task-relevant. This could occur because all three contexts are treated as individual contexts throughout the whole experiment (e.g., Cañadas et al., 2013). Or it could occur because associations formed in the first block interfere with learning new associations in the second (e.g., Brosowsky \& Crump, 2016). On other hand, task-relevance might not only dictate how associations are formed, but also whether learned associations are used. In this case, we would expect smaller congruency effects when the high-conflict context dimension is made task-relevant as compared to when the low-conflict context dimension is made task-relevant. This would demonstrate that participants were able to ignore a previously learned association and adopt a new context-specific control strategy.

A secondary goal was to conceptually replicate previous findings and test the generalizability of any task-relevance effects. Therefore, we included three conditions that differed only in the kinds of stimuli used to create context cues. In one condition we used face images where context dimensions were defined by social categories, similar to Cañadas et al. (2013). However, they found that social categorization occurred spontaneously and it was unclear what impact that would have on the current results. Therefore, we included another condition that used images of simple objects where context dimensions were defined by object features. Finally, Cañadas et al. (2013) also found that context effects generalized to novel face images. In our third condition we used non-repeating face exemplars to examine whether any task-relevance effects would also generalize in our design or if learning would be image-specific.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

@@

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

All participants were recruited from Amazon Mechanical Turk (AMT) and compensated \$2.00 for participating. The amount compensated was calculated by estimating the maximum amount of time required to complete each experiment and multiplying by \$6.00 per hour. For each experiment the number of HITs (Human intelligence tasks, an Amazon term for a work-unit) refers to the number of participants who initiated the study. Participants were included in the study if they completed all trials. We posted 150 HITs and 144 participants completed all trials.

\hypertarget{apparatus-stimuli}{%
\subsection{Apparatus \& Stimuli}\label{apparatus-stimuli}}

@@

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/figure2-1.pdf}
\caption{\label{fig:figure2}Results from experiment 1 showing congruency effects (incongruent - congruent) for frequency unbiased items in reaction times (left) and error rates (right) as a function of the task-relevant context (100\% proportion congruent versus 0\% proportion congruent).}
\end{figure}

\hypertarget{results}{%
\section{Results}\label{results}}

@@

Similarly, the corresponding error analysis also resulted in a significant two-way interaction between the task-relevant context and unbiased-item congruency, \(F(1, 128) = 6.72\), \(\mathit{MSE} = 13.93\), \(p = .011\), \(\hat{\eta}^2_p = .050\), and a non-significant three-way interaction between task-relevant context, unbiased-item congruency, and context-type, \(F(2, 128) = 0.92\), \(\mathit{MSE} = 13.93\), \(p = .400\), \(\hat{\eta}^2_p = .014\). Therefore, the congruency effects in error rates were also smaller when the context dimension associated with high conflict was made task-relevant, corroborating the results of the reaction time analysis.

\begingroup
\setlength{\parindent}{0in}
\vspace{2.5mm}

\textbf{Table 1. \\ \textit{Reaction times and error rates from Experiment 1.}}

\includegraphics{images/table1.pdf}
\endgroup

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

The aim of the current study was to determine whether task-relevance plays a role in the contextual recruitment of selective attention. Specifically, we tested whether manipulating the relative task-relevance of context cues could cause participants to ignore a previously learned context-association and apply a new association. We created a frequency unbiased context cue that shared features with two frequency biased contexts and used a feature-counting task to manipulate the task-relevance of context dimensions across two blocks of trials. Critically, halfway through the experiment participants received new instructions changing the task-relevant feature from one frequency biased cue to the other.

@@

In light of prior work, we take this as evidence that the contextual recruitment of selective attention, although likely implicit, is not obligatory (e.g., Brosowsky \& Crump, 2016), requiring that environmental information be incorporated into the task representation. Similarly, there appears to be flexibility in which environmental features are selected and used to guide attention, which can be rapidly updated depending on the task-relevance of those features. Such a result lends some insight into how context-specific control might operate in more complex, real-world environments, where there is an over-abundance of environmental features that afford many different learned associations. From a theoretical perspective, this result is consistent memory-based accounts of CSPC phenomena (Brosowsky \& Crump, 2018; Bugg \& Hutchison, 2013; Crump et al., 2017, 2018; Crump \& Milliken, 2009). Under this view, a memory process encodes attentional priorities in the representation of individual experiences and, as a result, become associated with the environment where they were used. The subsequent reoccurrence of a prior context triggers the retrieval and reinstatement of those attentional priorities. Our results show however, that all the features of the environment may not be treated equally and that only task-relevant features are used to probe memory and guide attention.

@@

\hypertarget{references}{%
\section{References}\label{references}}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-abrahamse_grounding_2016}{}%
Abrahamse, E., Braem, S., Notebaert, W., \& Verguts, T. (2016). Grounding cognitive control in associative learning. \emph{Psychological Bulletin}, \emph{142}, 693--728. doi:\href{https://doi.org/10.1037/bul0000047}{10.1037/bul0000047}

\leavevmode\hypertarget{ref-alards-tomalin_auditory_2017}{}%
Alards-Tomalin, D., Brosowsky, N. P., \& Mondor, T. A. (2017). Auditory statistical learning: Predictive frequency information affects the deployment of contextually mediated attentional resources on perceptual tasks. \emph{Journal of Cognitive Psychology}, \emph{0}(0), 1--11. doi:\href{https://doi.org/10.1080/20445911.2017.1353518}{10.1080/20445911.2017.1353518}

\leavevmode\hypertarget{ref-blais_trial-by-trial_2015}{}%
Blais, C., Harris, M. B., Sinanian, M. H., \& Bunge, S. A. (2015). Trial-by-trial adjustments in control triggered by incidentally encoded semantic cues. \emph{The Quarterly Journal of Experimental Psychology}, \emph{68}, 1920--1930. doi:\href{https://doi.org/10.1080/17470218.2014.1000346}{10.1080/17470218.2014.1000346}

\leavevmode\hypertarget{ref-brady_visual_2013}{}%
Brady, T. F., Konkle, T., Gill, J., Oliva, A., \& Alvarez, G. A. (2013). Visual long-term memory has the same limit on fidelity as visual working memory. \emph{Psychological Science}, \emph{24}, 981--990. doi:\href{https://doi.org/10.1177/0956797612465439}{10.1177/0956797612465439}

\leavevmode\hypertarget{ref-brewer_dual_1988}{}%
Brewer, M. B. (1988). A dual process model of impression formation. In R. S. Wyer Jr \& T. K. Srull (Eds.), \emph{A dual-process model of impression formation: Advances in social cognition} (Vol. 1, pp. 1--36). Hillsdale, NJ: Erlbaum.

\leavevmode\hypertarget{ref-brosowsky_context-specific_2016}{}%
Brosowsky, N. P., \& Crump, M. J. (2016). Context-specific attentional sampling: Intentional control as a pre-requisite for contextual control. \emph{Consciousness and Cognition}, \emph{44}, 146--160. doi:\href{https://doi.org/10.1016/j.concog.2016.07.001}{10.1016/j.concog.2016.07.001}

\leavevmode\hypertarget{ref-brosowsky_memory-guided_2018}{}%
Brosowsky, N. P., \& Crump, M. J. C. (2018). Memory-guided selective attention: Single experiences with conflict have long-lasting effects on cognitive control. \emph{Journal of Experimental Psychology: General}. doi:\href{https://doi.org/10.1037/xge0000431}{10.1037/xge0000431}

\leavevmode\hypertarget{ref-bugg_conflict-triggered_2014}{}%
Bugg, J. (2014). Conflict-triggered top-down control: Default mode, last resort, or no such thing? \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{40}(2), 567.

\leavevmode\hypertarget{ref-bugg_support_2012}{}%
Bugg, J., \& Crump, M. J. (2012). In support of a distinction between voluntary and stimulus-driven control: A review of the literature on proportion congruent effects. \emph{Frontiers in Psychology}, \emph{3}, 367. doi:\href{https://doi.org/10.3389/fpsyg.2012.00367}{10.3389/fpsyg.2012.00367}

\leavevmode\hypertarget{ref-bugg_converging_2013}{}%
Bugg, J., \& Hutchison, K. A. (2013). Converging evidence for control of color--word Stroop interference at the item level. \emph{Journal of Experimental Psychology: Human Perception and Performance}, \emph{39}(2), 433.

\leavevmode\hypertarget{ref-bugg_multiple_2008}{}%
Bugg, J., Jacoby, L. L., \& Toth, J. P. (2008). Multiple levels of control in the Stroop task. \emph{Memory \& Cognition}, \emph{36}, 1484--1494. doi:\href{https://doi.org/10.3758/MC.36.8.1484}{10.3758/MC.36.8.1484}

\leavevmode\hypertarget{ref-canadas_social_2013}{}%
Cañadas, E., Rodríguez-Bailón, R., Milliken, B., \& Lupiáñez, J. (2013). Social categories as a context for the allocation of attentional control. \emph{Journal of Experimental Psychology: General}, \emph{142}, 934--943. doi:\href{https://doi.org/10.1037/a0029794}{10.1037/a0029794}

\leavevmode\hypertarget{ref-corballis_independent_2003}{}%
Corballis, P. M., \& Gratton, G. (2003). Independent control of processing strategies for different locations in the visual field. \emph{Biological Psychology}, \emph{64}, 191--209. doi:\href{https://doi.org/10.1016/S0301-0511(03)00109-1}{10.1016/S0301-0511(03)00109-1}
